Nail: A practical interface generator for data formats

### 2014-04-30 paper/osdi14/relwk.tex
Related work

\paragraph{Motivation.}
\XXX[Move to the intro, into its own section, or kill?]

@@

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.
By allowing developers to specify their data format just once, Nail
avoids logic errors and inconsistencies in parsing and output generation.

@@

\XXX[Parsifal at langsec '14]
\paragraph{Related Work}
Proper input recognition has  been shown to be an excellent way of
eliminating malicious inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.

Parser generators for binary protocols were first introduced by the
Hammer~\cite{hammer-parser} parser.
While previous parser generators could also be used to write grammars
for binary protocols,\footnote{Theoretically speaking, the alphabet
over which a grammar is an  abstract set, so most algorithms work
just as well on an alphabet of $\{0,1\}$.} doing so is practically
inconvenient. Hammer allows the programmer to specify a grammar in
terms of bits and bytes instead of characters. Common concerns, such as
endianness and bit-packing are transparently hidden by the library. Hammer
implements grammars as language-integrated parser combinators, an approach
popularized by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
parser combinator style (to our knowledge, first  described
in~\cite{burge1975recursive}) is a natural way of concisely expressing
top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
more background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the default
parsing algorithm used by Hammer.} by composing them from one or multiple
sub-parsers.  Hammer then constructs a tree of function pointers which
can be invoked to parse a given input into an abstract syntax tree (AST).

\paragraph{Application use of parsers.}
Generated parsers have long been used to parse human input, such as programming languages and
configuration files. Frequently, such languages are often specified with a formal grammar in an
executable form.

Unfortunately,  parser frameworks are seldom used  to recognize
machine-created input. For example, the security-critical and
well-engineered MIT Kerberos distribution uses parser generators, but
only for handling configuration files. A notable exception is the Mongrel
web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
for HTTP written in the Ragel~\cite{ragel-paper} regular expression
language. Mongrel was re-written from scratch multiple times to achieve
better scalability and design, yet the grammar could be  re-used across
all iterations~\cite{patterson-citation}.

>>>>

% \XXX[Move to the intro, into its own section, or kill?]

\XXX[Parsifal at langsec '14]
\paragraph{Parsers}

Parser generators for binary protocols were first introduced by the
Hammer~\cite{hammer-parser} parser.
While previous parser generators could also be used to write grammars
for binary protocols,\footnote{Theoretically speaking, the alphabet
over which a grammar is an  abstract set, so most algorithms work
just as well on an alphabet of $\{0,1\}$.} doing so is practically
inconvenient. Hammer allows the programmer to specify a grammar in
terms of bits and bytes instead of characters. Common concerns, such as
endianness and bit-packing are transparently hidden by the library.

Hammer implements grammars as language-integrated parser combinators, an approach popularized by
Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
described in~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of
  expressing grammars, see Bryan Ford's masters thesis~\cite{ford2002packrat}, which also describes
  the default parsing algorithm used by Hammer.} by composing them from one or multiple sub-parsers.
Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
an abstract syntax tree (AST).

\paragraph{Application use of parsers.}
Generated parsers have long been used to parse human input, such as programming languages and
configuration files. Frequently, such languages are often specified with a formal grammar in an
executable form. Unfortunately,  parser frameworks are seldom used  to recognize
machine-created input.

A notable exception is the Mongrel
web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
for HTTP written in the Ragel~\cite{ragel-paper} regular expression
language. Mongrel was re-written from scratch multiple times to achieve
better scalability and design, yet the grammar could be  re-used across
all iterations~\cite{patterson-citation}.

\paragraph{Other vulnerability mitigations.}

@@

The second class is logic errors, where application code misinterprets input data. Safe languages
and exploit mitigation technologies do not help against such vulnerabilities. This can lead to
serious security consequences when two systems disagree on the meaning of a network packet or a
signed message, as in iOS\footnote{The XNU kernel and the user code-signing verifier interpret
  executable metadata differently, so the code signature sees different bytes at a virtual address
  than the executable that runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing
and even the X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Logic errors are often
causing memory corruption bugs as well, for example when a protocol stores the size of a buffer in
two places and the implementation does not check their consistency. These mistakes are
highly application-specific, and are difficult to mitigate using existing techniques, and these
mistakes can occur even in high-level languages that guarantee memory safety. By allowing developers
to specify their data format just once, Nail avoids logic errors and inconsistencies in parsing and
output generation.

@@

Proper input recognition has  been shown to be an excellent way of
eliminating ambiguous inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.
